# Reproducible Research: Peer Assessment 1
My Assignment
========================

This is the R markdown file for the first Assignment.

##Loading and Preprocessing the Data

First, we're going to load some data.

```{r echo=TRUE}
data <- read.csv('activity.csv')
```

Then we will take a look at the data to see its columns

```{r echo=TRUE}
head(data)
```

## What is the mean total number of steps taken per day?
First, we calculate the cumulative sum of steps taken per day. For this, we use the data.table package.

```{r echo=TRUE}

library(data.table)
dat <- data.table(data)
dat[!is.na(steps), Cumulative.Sum := cumsum(steps), by = c('date')]
sumStepsDaily <- dat$Cumulative.Sum

```

Now, we make a histogram of the total number of steps taken each day.
```{r echo=TRUE}
hist(sumStepsDaily)
``` 

The mean total number of steps taken each day is:
```{r echo=TRUE}
mean(sumStepsDaily, na.rm=TRUE)
``` 

The median total number of steps taken each day is:
```{r echo=TRUE}
median(sumStepsDaily, na.rm=TRUE)
``` 

## What is the average daily activity pattern?

First, we calculate the average number of steps taken per time interval, averaged across all days, and store it in 'avgStepsInterval':
```{r echo=TRUE}
dat[!is.na(steps), Cumulative.Sum.Interval := cumsum(steps), by = c('interval')]
avgStepsInterval <- dat$Cumulative.Sum.Interval
``` 

Then, we Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE}
plot(data$interval, avgStepsInterval, type="l", main="Time Series Plot", sub="Average Steps per 5-minute Interval", xlab="5-minute interval", ylab="Average No. of steps taken (across all days)")
``` 

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

In order to find this out, we first find the largest average number of steps across all time intervals:
```{r echo=TRUE}
max(avgStepsInterval, na.rm=TRUE)
``` 

Then, we find out which row this value belongs to, and we see the value of the 5-minute interval which, on average, contains the maximum number of steps:
```{r echo=TRUE}
dat[which.max(avgStepsInterval),]
``` 
## Imputing missing values

# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

We do this as follows:
```{r echo=TRUE}
colSums(is.na(data))
``` 

My strategy for filling the NA values would be to replace them with 0, as nothing is known about this data.
```{r echo=TRUE}
dat$steps <- replace(dat$steps, is.na(dat$steps), 0) 
```

